The \dealii Library, Version 9.0

### 2018-05-14 9.0/paper.tex
Some last edits.

\item New caching mechanism for expensive grid computations: we introduced a new
  class \texttt{GridTools::Cache} that caches computationally intensive
  information about a \texttt{Triangulation}.

This class allows the user to query some of the data structures constructed
using functions in the \texttt{GridTools} namespace. This data is then
computed only once, and
then cached inside this class for faster access whenever the triangulation has
not changed. The cache is marked for update by the \texttt{Triangulation} itself
using signals.

@@

\dealii has had a tutorial program (step-33) since 2007 that
demonstrates this technique based on the Trilinos Sacado \cite{Bartlett2006a} package, but
the functionality was not available pervasively throughout
deal.II.
This has changed with release 9.0 where support is given for differentiation
using a selection of ``white-listed'' libraries (namely ADOL-C \citep{Griewank1996a}
and Sacado) and a subset of their supported number types. Currently,
we offer support for the following cases:
\begin{itemize}
\item ADOL-C taped (n-differentiable),
\item ADOL-C tapeless (once differentiable),
\item Sacado dynamic forward (once differentiable),
\item Sacado reverse (once differentiable),
\item Sacado nested dynamic forward (twice differentiable), and
\item Sacado nested reverse and dynamic forward (twice differentiable).
\end{itemize}
In practice, this support means that these ADOL-C and Sacado data
types can be used in the \texttt{FEValues}, \texttt{FEValuesViews},
\texttt{Tensor}, \texttt{SymmetricTensor},
and related classes that are generally used to assemble linear systems
and right hand sides.
Given the updated capabilities of the library, there is now a dedicated module
that presents the AD compatibility and capabilities of the \dealii libraries.
Furthermore, the use of Sacado is now  demonstrated in a much more simplified
and transparent manner in a modernized version of an existing ``code gallery''
example \cite{Pelteret2016a}.

@@

  \dealii now has interfaces to SUNDIALS's ARKode, IDA, and KINSOL sub-packages.
  ARKode is a solver library that provides adaptive-step time
  integration. IDA is a package for the solution of differential-algebraic
  equations systems in the form $F(t,y,y')=0$. KINSOL is solver for nonlinear
  algebraic systems.

@@

Finally, we have changed the entire code base to avoid using raw
pointers and instead use \texttt{std::unique\_ptr} and
\texttt{std::shared\_ptr} where possible to make memory management
more reliable. These changes include some minor incompatibilities: all
\texttt{clone()} functions (such as \texttt{FiniteElement::clone()} and
\texttt{Mapping::clone()}) now return \texttt{std::unique\_ptr}s instead of
C-style raw pointers. Indeed, nearly all interfaces throughout the library that return a pointer
now return either a \texttt{std::shared\_ptr} or a \texttt{std::unique\_ptr},
thereby clarifying object ownership responsibilities and avoiding memory leaks.

@@

This release of \dealii adds support for GPUs both for matrix-based and
matrix-free applications. For matrix-based applications, we rely on
cuSPARSE\cite{cusparse} and cuSOLVER\cite{cusolver} for operations on sparse
matrices such as matrix-vector multiplication and for direct solvers. We have
introduced a new type of sparse matrix, \texttt{CUDAWrappers::SparseMatrix},
which moves onto the device a \dealii \texttt{SparseMatrix} and changes the
format of the underlying data to the appropriate CSR format used by cuSPARSE. We
also have added wrappers for Cholesky and LU factorizations provided by
cuSOLVER. In practice, a user would assemble the matrix associated to the system
on the host and then move the matrix to the device. At this point, the system
would be solved on the device and the solution would be moved back to the host.

We also have some support for matrix-free computation on a GPU. For now, the
evaluation of the operator is limited to mesh without hanging-nodes.

@@

The tutorial program \texttt{step-60} shows how to perform computations on
non-matching grids, and it presents advanced manipulation of
\texttt{ParameterHandler} objects, exploiting the new \texttt{ParameterAccessor}
and \texttt{ParameterAccessorProxy} classes. \texttt{step-60} solves a Poisson
problem on a domain $\Omega$, subject to equality constraints defined on an
embedded domain $\Gamma$. The embedded domain can be of co-dimension one or
co-dimension zero, and its definition is independent with respect to $\Omega$.
In order to enforce correctly the constraints, a non-matching coupling matrix
needs to be constructed. This is achieved using the new
\texttt{NonMatching::create\_coupling\_sparsity\_pattern} and
\texttt{NonMatching::create\_coupling\_mass\_matrix}, that exploit some new
functionality in the \texttt{GridTools} namespace.

>>>>

\item A new caching mechanism for expensive grid computations: this version introduces a new
  class \texttt{GridTools::Cache} that caches computationally intensive
  information about a \texttt{Triangulation}.

This class allows the user to query some of the data structures constructed
using functions in the \texttt{GridTools} namespace. This data is then
computed only once, and
cached inside this class for faster access whenever the triangulation has
not changed. The cache is marked for update by the \texttt{Triangulation} itself
using signals so that data is properly invalidated upon mesh
refinement and coarsening.

@@

\dealii has had a tutorial program (step-33) since 2007 that
demonstrates this technique based on the Trilinos Sacado \cite{Bartlett2006a} package, but
the functionality was not available pervasively throughout
deal.II.
This has changed with release 9.0 where support for differentiation is
now available
using a selection of ``white-listed'' libraries (namely ADOL-C \citep{Griewank1996a}
and Sacado) and a subset of their supported number types. Currently,
we offer support for the following cases:
\begin{itemize}
\item ADOL-C taped (n-differentiable),
\item ADOL-C tapeless (once differentiable),
\item Sacado dynamic forward (once differentiable),
\item Sacado reverse (once differentiable),
\item Sacado nested dynamic forward (twice differentiable), and
\item Sacado nested reverse and dynamic forward (twice differentiable).
\end{itemize}
In practice, this support means that these ADOL-C and Sacado data
types can be used as the underlying ``scalar'' in the \texttt{FEValues}, \texttt{FEValuesViews},
\texttt{Tensor}, \texttt{SymmetricTensor},
and related classes that are generally used to assemble linear systems
and right hand sides.
Given the updated capabilities of the library, there is now a dedicated module
that presents the AD compatibility and capabilities of the \dealii libraries.
Furthermore, the use of Sacado is demonstrated in a much more simplified
and transparent manner in a modernized version of an existing ``code gallery''
example \cite{Pelteret2016a}.

@@

  \dealii now has interfaces to SUNDIALS's ARKode, IDA, and KINSOL sub-packages.
  ARKode is a solver library that provides adaptive-step time
  integration. IDA is a package for the solution of differential-algebraic
  equations systems in the form $F(t,y,y')=0$. KINSOL is a solver for nonlinear
  algebraic systems.

@@

Finally, we have changed the entire code base to avoid using raw
pointers and instead use \texttt{std::unique\_ptr} and
\texttt{std::shared\_ptr} where possible to make memory management
more reliable. These changes include some minor incompatibilities: all
\texttt{clone()} functions (such as \texttt{FiniteElement::clone()} and
\texttt{Mapping::clone()}) now return \texttt{std::unique\_ptr}s instead of
C-style raw pointers. Indeed, nearly all interfaces throughout the library that return a pointer
now return either a \texttt{std::shared\_ptr} or a \texttt{std::unique\_ptr},
thereby clarifying object ownership and avoiding memory leaks.

@@

This release of \dealii adds support for GPUs both for matrix-based and
matrix-free applications. For matrix-based applications, we rely on
cuSPARSE \cite{cusparse} and cuSOLVER \cite{cusolver} for operations on sparse
matrices such as matrix-vector multiplication and for direct solvers. We have
introduced a new type of sparse matrix, \texttt{CUDAWrappers::SparseMatrix},
which moves onto the device a \dealii \texttt{SparseMatrix} and changes the
format of the underlying data to the appropriate CSR format used by cuSPARSE. We
also have added wrappers for Cholesky and LU factorizations provided by
cuSOLVER. In practice, a user would assemble the system matrix and
right hand side vector
on the host and then move them to the device. At this point, the
linear system
would be solved on the device and the solution would be moved back to the host.

We also have some support for matrix-free computation on a GPU. For now, the
evaluation of the operator is limited to meshes without hanging nodes.

@@

The tutorial program \texttt{step-60} shows how to perform computations on
non-matching grids, and it presents advanced manipulation of
\texttt{ParameterHandler} objects using the new \texttt{ParameterAccessor}
and \texttt{ParameterAccessorProxy} classes. \texttt{step-60} solves a Poisson
problem on a domain $\Omega$, subject to equality constraints defined on an
embedded domain $\Gamma$. The embedded domain can be of co-dimension one or
co-dimension zero, and its definition is independent with respect to $\Omega$.
In order to enforce correctly the constraints, a non-matching coupling matrix
needs to be constructed. This is achieved using the new
\texttt{NonMatching::create\_coupling\_sparsity\_pattern} and
\texttt{NonMatching::create\_coupling\_mass\_matrix} functions that exploit new
functionality in the \texttt{GridTools} namespace.
